### 2011-02-24 trunk/deliverables/D2_4_Mediation/fast_d2_4_3_part1.tex
new files for version 2.4.3
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1282 d8b0c59b-6544-0410-94d9-83a5658a058e

To support manual ontology mapping, we plan to develop a rule based, interactive Facts Transformation Tool. This tool utilizes techniques of our service wrapper tool, cf. deliverable D4.3.2. One part of the service wrapper tool deals with a rule based approach for the transformation of server responses into FAST facts. This problem is closely related to the problem of ontology mediation. Thus, these techniques will also allow to create data transformation operators, i.e. operators that mediate between two ontologies.

Figure \ref{fig:FactsTransformationToolPortDefinitions} shows a photoshoped preview of the planned Facts Transformation Tool. The Facts Transformation Tool has two tabs. The \textit{General} tab allows to define the input and output ports for the desired data transformation operator. These ports shall be  typed according to the participating semantic elements to be derived from the FAST catalogue. This typing will provide the structure of the input and output data of the desired data transformation operator.

Figure \ref{fig:FactsTransformationToolTransformationRules} shows the planned editor for transformation rules. On the left of Figure \ref{fig:FactsTransformationToolTransformationRules} there is a tree of example data that is used to validate the transformation rules during editing. Thus, the Facts Transformation Tool will execute the transformation rules after any editing in order to give direct feedback on their functionality and to enable the user to validate the achieved transformation. The transformation results are shown on the right of Figure \ref{fig:FactsTransformationToolTransformationRules}.

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{images/FactsTransformationToolGVSWithPortDefinitionsNoGlitch2.png}
    \caption{Definition of operator pre- / postconditions (photoshoped screen dump)}
    \label{fig:FactsTransformationToolPortDefinitions}
\end{figure}

@@

Note that simple one-to-one transformation rules should be derived with the help of the ontology mediation approach described in Section \ref{subsubsec:automaticmapping}. Such simple rules should be added to the transformation tool, automatically. The more complex rules may then be added interactively. During editing, the semantics provided by the FAST catalogue will be used to support the user with completion proposals and consistency checking. In addition, the direct execution of the rules on example data will help the user to create appropriate rules and to validate their effects.

So far our transformation rules are restricted to simple cases. Thus, we have focused on simple cases allowing a simple transformation approach that does not require sophisticated skills from our users. However, as discussed on the Amazon to GoodRelations example, there exist more complex cases. In the third year of the project we will investigate how such more complex cases can be covered, too, and whether it is still possible to keep the rules simple and understandable.

@@

Once the data transformation operator has been defined, it will be added to the FAST catalogue in order to allow its use for the definition of piping configurations of new screens within the FAST CGS tool. For this purpose a simple semantic description will be sent to FAST catalogue defining the semantics of the operator and the semantics of its input and output ports.

Currently, the Service Wrapper Tool uses its own repository for trees of transformation rules. First of all this has technical reasons: the development of the transformation tool is based on a reference architecture and libraries for web based interactive systems provided by Kassel University, cf. \cite{AsDH2009CVSM}. This repository has the special advantage that it provides support for teams of developers, thus the service wrapper tool allows to merge the contributions of multiple developers. However, on the long run, we shall store the transformation rules together with the service wrapper and transformation operator definitions in the FAST catalogue.

@@

At the gadget building phase, whenever preconditions and postconditions of screens, resources or operators show a semantic mismatch, the catalogue will offer the possible matching operators that would solve the incompatibility problem. This way a matching operator can be placed between two pre- postconditions that would otherwise be incompatible, making interoperation between the two possible. In seldom cases, the catalogue may fail to retrieve an appropriate mediation operator. In such cases, the possibility of building a new matching operator with the help of the Facts Transformation Tool will be presented, in order to meet the user's need.

\section{Runtime} 
\label{sec:runtime}
For deployment, the gadgets implementation is generated in Javascript. In addition, a small FAST runtime library (again Javascript) is added to the gadget. This runtime library provides notification mechanisms and manages the fact base of the gadget. For data transformation we will generate Javascript code that connects itself to the fact base and any time a fact is sent to the input port of the operator, it will do the data transformation and send the result to the output port. The code generated from the data transformation rules follows a recursive pattern where each rule becomes a Javascript function with parameters for the source and target fact subtrees to be transformed. The function body then uses code to search the source tree for sub elements that match the \textit{From} part of the rule and then code for the desired operation is generated. For debugging purposes the user may examine the code generated for an operator in the \textit{CodeGen Viewer} tab of the Facts Transformation tool.

>>>>

To support manual ontology mapping, we have developed a rule based, interactive Facts Transformation Tool. This tool utilizes techniques of our service wrapper tool, cf. deliverable D4.3.2. One part of the service wrapper tool deals with a rule based approach for the transformation of server responses into FAST facts. This problem is closely related to the problem of ontology mediation. Thus, these techniques also allow to create data transformation operators, i.e. operators that mediate between two ontologies.

Figure \ref{fig:FactsTransformationToolPortDefinitions} shows a screen shot of the  Facts Transformation Tool within the GVS environment. The Facts Transformation Tool has three tabs. The \textit{General} tab allows to define the pre- and postconditions for the desired data transformation operator. These conditions are  typed according to the participating semantic elements to be derived from the FAST catalogue. This typing will provide the structure of the input and output data of the desired data transformation operator.

Figure \ref{fig:FactsTransformationToolTransformationRules} shows the editor for transformation rules. On the left of Figure \ref{fig:FactsTransformationToolTransformationRules} there is a tree of example data that is used to validate the transformation rules during editing. Thus, the Facts Transformation Tool will execute the transformation rules after any editing in order to give direct feedback on their functionality and to enable the user to validate the achieved transformation. The transformation results are shown on the right of Figure \ref{fig:FactsTransformationToolTransformationRules}.

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{images/FactsTransformationToolGVSWithPortDefinitionsNoGlitch2.png}
    \caption{Definition of operator pre- / postconditions}
    \label{fig:FactsTransformationToolPortDefinitions}
\end{figure}

@@

Note that simple one-to-one transformation rules may be derived with the help of the ontology mediation approach described in Section \ref{subsubsec:automaticmapping}. Such simple rules may be added to the transformation tool, automatically. We try to add this functionality within the remaining project time. The more complex rules may then be added interactively. During editing, the semantics provided by the FAST catalogue is used to support the user with completion proposals and consistency checking. In addition, the direct execution of the rules on example data helps the user to create appropriate rules and to validate their effects.

So far our transformation rules are restricted to simple cases. We have focused on simple cases allowing a simple transformation approach that does not require sophisticated skills from our users.

@@

Once the data transformation operator has been defined, it may be added to the FAST catalogue in order to allow its use for the definition of piping configurations of new screens within the FAST CGS tool. For this purpose a simple semantic description is sent to the FAST catalogue defining the semantics of the operator and the semantics of its pre- and postconditions.

The Service Wrapper Tool uses its own repository for trees of transformation rules. First of all this has technical reasons: the development of the transformation tool is based on a reference architecture and libraries for web based interactive systems provided by Kassel University, cf. \cite{AsDH2009CVSM}. This repository has the special advantage that it provides support for teams of developers, thus the service wrapper tool allows to merge the contributions of multiple developers.

@@

At the gadget building phase, whenever preconditions and postconditions of screens, resources or operators show a semantic mismatch, the catalogue offers the possible matching operators that would solve the incompatibility problem. This way a matching operator can be placed between two pre- postconditions that would otherwise be incompatible, making interoperation between the two possible. In seldom cases, the catalogue may fail to retrieve an appropriate mediation operator. In such cases, the possibility of building a new matching operator with the help of the Facts Transformation Tool is presented, in order to meet the user's need.

\section{Runtime} 
\label{sec:runtime}
For deployment, the gadgets implementation is generated in Javascript. In addition, a small FAST runtime library (again Javascript) is added to the gadget. This runtime library provides notification mechanisms and manages the fact base of the gadget. For data transformation we generate Javascript code that connects itself to the fact base and any time a fact is sent to the precondition of the operator, it does the data transformation and sends the result to the postcondition. The code generated from the data transformation rules follows a recursive pattern where each rule becomes a Javascript function with parameters for the source and target fact subtrees to be transformed. The function body then uses code to search the source tree for sub elements that match the \textit{From} part of the rule and then code for the desired operation is generated. For debugging purposes the user may examine the code generated for an operator in the \textit{CodeGen Viewer} tab of the Facts Transformation tool.

### 2011-02-24 trunk/deliverables/D2_4_Mediation/fast_d2_4_3_part2.tex
new files for version 2.4.3
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1282 d8b0c59b-6544-0410-94d9-83a5658a058e

\begin{itemize}
\item \textit{instance-based/schema-based} matching: A schema-based
matcher uses different approaches to determine correspondence between
ontologies based on concepts and relations \cite{debruin2005wsml}. 
An instance-based matcher takes instances
(i.e. instance data) belonging to the different concepts in the
ontologies and uses these to discover similarities between the concept.

>>>>

\begin{itemize}
\item \textit{instance-based/schema-based} matching: A schema-based
matcher uses different  \linebreak approaches to determine correspondence between
ontologies based on concepts and relations \cite{debruin2005wsml}. 
An instance-based matcher takes instances
(i.e. instance data) belonging to the different concepts in the
ontologies and uses these to discover similarities between the concept.
